---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# assemblerr

assemblerr is an R package to construct pharmacometric models by combining pre-defined model components. It's intended to simplify the specification of complex pharmacometric models and provide a mechanism to generate models in an automatic way. With assemblerr, models are specified using R code and then converted to a NONMEM control stream.   

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("UUPharmacometrics/assemblerr")
```
## Example

This is a basic example which shows you how to create an Emax model and generate the corresponding NONMEM code:

```{r example}
library(assemblerr)

m <- model() +
  prm_log_normal("emax") +
  prm_normal("ed50") +
  obs_additive(effect~emax*dose/(ed50+dose)) 

render(m) 

```

