---
title: "Experimental"
author: "Sebastian Ueckert"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Experimental}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette documents some of the more experimental features of the package.

# PRED models

assemblerr autmoatically generates a PRED model when no differential equations are present.


```{r}
library(assemblerr, warn.conflicts = F)

# construct one-compartment 1st order elimination model
m <- model()+
  algebraic(effect~emax*dose/(ed50+dose))+
  prm_log_normal("EMAX")+
  prm_log_normal("ED50")+
  obs_additive(eff~effect, "effect")+
  obs_additive(~emax, "emax")

# convert to a NONMEM model and generate NMTran code
as_nm_model(m)  %>% 
  render %>%
  cat()
```

# Automatic ordering of variables

assemblerrr uses a declarative philosophy to specify models, in consequence the order of statements does not matter. When the NONMEM model is generated, assemblerr will determine the appropriate order of instructions and assign each instruction to the suitable block (PK, DES, ERROR) . 

```{r}


m <- model()+
  compartment("central", volume = ~vc) +
  flow(from = "central", definition = ~cl*C) +
  flow(to = "central", definition = ~kin) +
  prm_log_normal("cl") +
  prm_log_normal("vc") +
  algebraic(k~cl/v)+
  algebraic(conc~A["central"])+
  algebraic(kin~k*A["central"])+
  obs_additive(~conc, "conc") +
  parameter_value_table(c(cl = 4.5, vc = 1), types = "typical") +
  parameter_value_table(c(cl = 0.09, vc = 0.09), types = "iiv")+
  parameter_value("ruv-add", "ruv", 1)


# convert to a NONMEM model and generate NMTran code
as_nm_model(m)  %>% 
  render %>%
  cat()

```

# Multiple observation models

assemblerr supports multiple observation models per model

```{r, eval=F}
m <- model()+
  compartment("central", volume = ~vc) +
  flow(from = "central", definition = ~cl*C) +
  prm_log_normal("cl") +
  prm_log_normal("vc") +
  prm_normal("emax")+
  prm_novar("ed50")+
  algebraic(conc~C["central"])+
  obs_additive(~conc, name = "PK") +
  obs_additive(~emax*conc/(conc+ed50), name = "PD")+
  parameter_value_table(c(cl = 4.5, vc = 1), types = "typical") +
  parameter_value_table(c(cl = 0.09, vc = 0.09), types = "iiv")+
  parameter_value("ruv-add", "ruv", 1)

m %>% 
  as_model_nm() %>% 
  render() %>% 
  cat()

```

# Model visualization

assemblerr can create a simple visualization of the model

```{r, fig.height=5, fig.width=7}
pkm <- pk_model()+
  pk_distribution_3cmp()+
  pk_elemination_linear()+
  pk_absorption_rate_fo()

pkm %>% 
  as_model() %>% 
  visualize()

```

# Conversion diagram

```{r, echo=FALSE, fig.height=5, fig.width=7}
DiagrammeR::mermaid("
graph TD
subgraph domain specific
A[pk_model]
B[pbpk_model]
C[pd_model] 
end
A --> D[model]
B --> D
C --> D
subgraph software specifc
E[model_nm]
F[model_mlx]
G[model_poped]
end
D --> E
D --> F
D --> G
")
```

# Conditional declarations

assemblerr supports conditional declarations through the function `cases`. Examples

- Different typical values for males and females `cl~cases(sex, 0 ~ theta[1], 1 ~ theta[2])*exp(eta[1])`
- Probabilities for binary response `p~cases(y, 0 ~ p0, 1 ~ 1 - p0)`
- Censoring due to quantification limits `cases(conc, conc > 1 ~ ipred, conc <= 1 ~ prob)`


```{r}
as_declaration(cl~cases(sex, 0 ~ theta[1], 1 ~ theta[2])*exp(eta[1])) %>% 
  render(render_opts_nm()) %>% 
  cat()
```

# Binary observations

```{r, eval =FALSE}
m <- model()+
  prm_logit("pbase")+
  obs_binary(~pbase, "effect")

# convert to a NONMEM model and generate NMTran code
as_nm_model(m)  %>% 
  render %>%
  cat()
```

# Support for PKPDsim

```{r, eval=FALSE}
m <- model()+
  compartment("central", volume = ~vc) +
  flow(from = "central", definition = ~cl*C) +
  prm_log_normal("cl") +
  prm_log_normal("vc")

m_sim <- m %>% 
  as_model_pkpd_sim()

m_sim$odes

```

