---
title: "Introduction"
author: "Sebastian Ueckert"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


assemblerr is an R package to construct pharmacometric models by combining pre-defined model components. It's intended to simplify the specification of complex pharmacometric models and provide a mechanism to generate models in an automatic way. With assemblerr, models are specified using R code and then converted to a NONMEM control stream (in the future the support of different modeling software is planned).   

# Quick start

## Load assemblerr 

```{r}
library(assemblerr, warn.conflicts = F)
```

## Build an Emax dose-response model

```{r}
m <- model()+
  prm_log_normal("EMAX")+
  prm_log_normal("ED50")+
  obs_additive(effect~emax*dose/(ed50+dose))

m
```


## Generate NONMEM code

```{r}
m %>% 
  as_nm_model() %>% 
  render() %>% 
  cat()
```


# Workflow

assemblerr's workflow is similar to ggplot, where different components are added to a model 


# Low-level building blocks

```{r}
library(assemblerr, warn.conflicts = F)

# construct one-compartment 1st order elimination model
m <- model()+
  compartment("central", volume = ~vc) +
  flow(from = "central", definition = ~cl*C) +
  prm_log_normal("cl")+
  prm_log_normal("vc")+
  obs_additive(conc~C["central"])+
  parameter_values(cl = c(log_mu = 4.5, log_sigma = 0.1),
                   vc = c(log_mu = 1, log_sigma = 0.1))

# convert to a NONMEM model and generate NMTran code
as_nm_model(m)  %>% 
  render %>%
  cat()

```


# Pharmacokinetic models

For PK models assemblerr supports a higher level of abstraction to simplify the construction of these types of models.

```{r}
# construct one-compartment 1st order elimination model
pkm <- pk_model() +
  pk_elemination_linear() +
  pk_distribution_2cmp() +
    parameter_values(cl = c(typical = 4.5, iiv = 0.09),
                   vc = c(typical = 1, iiv = 0.09),
                   ruv = c(sd = 0.1)) +
  meta_tag("title", "run1")

# convert to a NONMEM model and generate NMTran code
pkm %>% 
  as_nm_model() %>% 
  render %>%
  cat()

```

The basic building blocks can still be added to a PK model, allowing to customize the pre-defined structure.

```{r}
# change the parameter model for clearance
pkm2 <- pkm +
  prm_normal("cl")

# convert to a NONMEM model and generate NMTran code
pkm2 %>% 
  as_nm_model() %>% 
  render %>%
  cat()

```

# Generating a range of models

A whole combination of models can be generated with the help of tools from the tidyverse. For example, one could do the following to generate one and two compartment models with linear and Michaelis-Menten elimination: 

```{r, eval=FALSE}
library(tidyverse, warn.conflicts = F)

# parameter names 
prm_names <- c("cl", "vc", "vp", "q", "km", "vmax")

# parameter values table with random theta values
pmt <- {
  runif(6, 0, 10) %>% 
    set_names(prm_names) %>% 
    parameter_value_table(type = "typical")} + 
    {rep(0.09, 6) %>% 
        set_names(prm_names) %>% 
      parameter_value_table(type = "iiv")} +
  parameter_value("ruv-add", "ruv", 1)

# create a list of all possible combinations
list(
     'distribution'=c("one-compartment", "two-compartment"),
     'elimination' = c("linear", "mm")) %>%
  cross() %>%
  # make components and add to models
  modify_depth(1, ~imap(.x, ~pk_component(.y, .x)) %>%
                 reduce(`+`, .init = pk_model())) %>%
  # add parameter table
  map(~.x+pmt) %>%
  # convert to NONMEM models
  map(as_model_nm) %>%
  # generate NMTran code
  map(render) %>% 
  # print code
  walk(cat)

```

Create all possible PK models and write to current working directory
```{r, eval=F}

list(
     'absorption-rate' = c("first-order","zero-order"),
     'absorption-delay' = c("none", "alag"),
     'distribution'=c("one-compartment", "two-compartment", "three-compartment"),
     'elimination' = c("linear", "mm")) %>%
  cross() %>%
  modify_depth(1, ~imap(.x, ~pk_component(.y, .x)) %>%
                 reduce(`+`, .init = pk_model())) %>%
  map(~as_model_nm(.x)) %>%
  imap(~write_model(.x, paste0("run", .y, ".mod")))
```
